---
title: "Sedentary"
subtitle: "I'm slowing down"
author: "Nick Cotter"
date: "2023-03-08"
categories: [exercise, analysis, exploration]
---

How much am I slowing down over time? And how much did the pandemic affect me? Let's take a look with an Apple Health export.

```{r libraries, warning=FALSE, echo=FALSE}
library(XML)
library(dplyr)
library(ggplot2)
library(tidyverse)
```

Unzip the export from Apple Health:

```{r cache=TRUE}
unzip('data/export-080323.zip')
```

Turn the export into a data frame:

```{r cache=TRUE}
health_xml <- xmlParse("apple_health_export/export.xml")
activity_df <- XML:::xmlAttrsToDataFrame(health_xml["//ActivitySummary"], stringsAsFactors = FALSE) %>%
as_tibble()
```

The column I'm interested in here is activeEnergyBurned - here is a plot:

```{r plot1, warning=FALSE}
plot(as.Date(activity_df$dateComponents), activity_df$activeEnergyBurned)
```

That looks a bit strange - what's with the date range?

```{r}
range(activity_df$dateComponents)
```

So a pre-epoch entry exists, which is clearly bogus. I can remove it - I'm primarily interested in the pre-pandemic period and follows:

```{r filtering, warning=FALSE}
activity <- activity_df %>%
  mutate(date = as.Date(dateComponents)) %>%
  mutate(calories = as.numeric(activeEnergyBurned)) %>%
  filter(date > "2010-1-1")
```

Let's plot it again:

```{r}
plot(activity$date, activity$calories)
```

Here is the trend line:

```{r trendline, warning=FALSE}
activity %>%
  ggplot(aes(x = date, y = calories)) + geom_line() + geom_smooth(formula = y ~ x, method = "loess") 
```

So, it's all been downhill since the end of 2019.
